The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-21 paper.tex
Cleaned up sec 5, moved some stuff to discussion, {phases->levels}

The so-called \emph{phase I} is the instinctive, pre-cognitive identification of
living peers. That humans tend to anthropomorphize robots intuitively in this
pre-cognitive way
is supported by studies done by Rosenthal-von der Pütten
\textit{et al.}~\cite{rosenthal-vonderputten_experimental_2013} who investigated
the neural correlates of emotional reactions of humans towards a robot. {\it
Empathy} is characteristic of this stage~\cite{rosenthalvonderPutten2013neural}.
Anthropomorphism at this pre-cognitive stage might also be mediated by the
human's mirror neurons system (neurons that fire both during execution of
specific goal-oriented action and during the viewing of action directed toward
the same goal~\cite{Rizzolatti1996, Kilner2009}) by allowing for a mapping of
the robot's goal-directed actions into the human own motor
repertoire~\cite{Gallese1998, Wolpert2003, cullen2013individual}.

After a longer observation period (typically including complete action sequences
of the robot) or short interaction (touching, short talk like greetings), we
suggest the human enters the cognitive \emph{phase II}: in this phase, the human
starts building a behavioral and cognitive model of the robot that would support
both the observed and imagined capabilities of the robot.  The \emph{familiarity
thesis}~\cite{hegel_understanding_2008} supports the idea that the human
first projects onto the robot mental models of similar agents he/she is already
familiar with (ranging from animals to human adults, to pets and children). We 
hypothesize that the nature of the projected mental
model, as well as how deep the human engages in this projection, might be
driven by the same parameters as we mentioned for the \emph{initial capital of
anthropomorphism} (section~\ref{sec:ica}).

The cognitive \emph{phase III} occurs after a \emph{contextualized} interaction.
A \emph{contextualized} interaction is \emph{explicitly purposeful} (the purpose
of the interaction, be it purely entertainment, is explicit and conscious to the
human), and takes place in an environment that fosters a stronger cognitive (and
possibly affective/social) commitment from the human in the interaction
(typically, at home). During this interaction, the human iteratively restates
and reshapes his/her behavioral and mental model of the robot (\emph{How does
the robot react to such and such situation/input?  What does the robot know
about me? About itself? About our environment? What can the robot learn?}, etc.).

@@

These cognitive phases overlap but do not exactly match the
\emph{Initialization}, \emph{Familiarization} and \emph{Stabilization} phases
introduced in our model of the dynamics of anthropomorphism. In particular,
cognitive phases I and II are both included in the \emph{initialization} phase
of the anthropomorphism model. Sub-cognitive anthropomorphism typically
\emph{initiates} the novelty effect by rapidly engaging the human in the
interaction through an initial projected agency, whereas cognitive phase II
(projection of familiar mental models) supports the novelty effect by inducing
beliefs that the robot is set up with possibly complex cognitive abilities.

The cognitive phase III also overlaps with the \emph{familiarization} phase: as
the human gets used to the robot, we hypothesize one restates and adapts its
cognitive model of the robot by iteratively reshaping pre-existent, familiar
models until it provides a satisfying support to explain and justify the
observed robot behavior.

A \emph{stable level of anthropomorphism} is reached when the adaptation process
depicted in cognitive phase III reached a stable state, \ie the user's
experience with the robot is correctly supported by the cognitive model he/she
has built.

\paragraph{Limits} This discussion on the cognitive correlates of the dynamics
of anthropomorphism are speculative, and only indirectly supported by
experimental evidence. New experiences need to be designed to specifically test
these hypothesis.\fxnote{...}

\subsection{Impact of the behavioral complexity of robots}

TBD. Most experiments are carried with robots exhibiting relatively simple
behaviors. In our model, the maximum level of anthropomorphism is reached with
the novelty effect, and then decrease. Can we go \emph{over} the novelty effect
if the robot exhibit more complex behaviors?

\subsection{Segmentation of interaction phases}

TBD

\subsection{Behavioral cues of anthropomorphism}
\label{sec:behavioralcues}

Our model suggests to analyze anthropomorphism based on a \emph{relative and
normalized level of anthropomorphic effects} instead of an absolute
quantitative assessment of anthropomorphic evidences. Similarly we introduce
two specific \emph{abstract} levels of anthropomorphism, the \emph{Initial
Capital of Anthropomorphism} and the \emph{Stabilized Level of
Anthropomorphism}: while we discuss the factors on which those values are
build, we did not formally investigated the experimental cues or evidences
that would allow to actually quantify those levels.

The intricacy and interdependency of those behaviors make them difficult to
quantify.  Their systematic identification and interpretation rely on
psychosomatic cues that the authors did not investigate.

A formal methodology for the assessment of anthropomorphic effects in HRI, as
well as a typology of these effects is still to be formally devised.

@@

@@

\begin{itemize}
	\item \textbf{Hypothesis 2} A person's tendency to anthropomorphize a robot decreases over time and with growing experience the person has with the robot. \textit{(familiarize oneself with the robot)}
\end{itemize}

@@

One related question we plan to elaborate on is the effects of \emph{mutual
explicitation of the mental model of the other agent}: a human and a robot
interact on a given task, and after a while, we interrupt the task and ask each
of the agents to explicit the mental model it has built of its partner
(emotional state, beliefs, intentions, etc.). Depending on the cognitive phase
the two agents are in, we may expect varying accuracy levels in the produced
mental models.

@@

@@

The article also discusses the cognitive correlates of anthropomorphism. We
proposes to identify three cognitive phases corresponding to successive
refinements of the mental models of the robot that the user builds during the
interaction. We show how these phases relate to observable anthropomorphic
effects, and how they evolve over time.

>>>>

The so-called \emph{level 1} is the instinctive, pre-cognitive identification of
living peers. That humans tend to anthropomorphize robots intuitively in this
pre-cognitive way
is supported by studies done by Rosenthal-von der Pütten
\textit{et al.}~\cite{rosenthal-vonderputten_experimental_2013} who investigated
the neural correlates of emotional reactions of humans towards a robot. {\it
Empathy} is characteristic of this stage~\cite{rosenthalvonderPutten2013neural}.
Anthropomorphism at this pre-cognitive stage might also be mediated by the
human's mirror neurons system (neurons that fire both during execution of
specific goal-oriented action and during the viewing of action directed toward
the same goal~\cite{Rizzolatti1996, Kilner2009}) by allowing for a mapping of
the robot's goal-directed actions into the human own motor
repertoire~\cite{Gallese1998, Wolpert2003, cullen2013individual}.

After a longer observation period (typically including complete action sequences
of the robot) or short interaction (touching, short talk like greetings), we
suggest the human reaches the cognitive \emph{level 2}: in this phase, the human
starts building a behavioral and cognitive model of the robot that would support
both the observed and imagined capabilities of the robot.  The \emph{familiarity
thesis}~\cite{hegel_understanding_2008} supports the idea that the human
first projects onto the robot mental models of similar agents he/she is already
familiar with (ranging from animals to human adults, to pets and children). We 
hypothesize that the nature of the projected mental
model, as well as how deep the human engages in this projection, might be
driven by the same parameters as we mentioned for the \emph{initial capital of
anthropomorphism} (section~\ref{sec:ica}).

The cognitive \emph{level 3} is reached after a \emph{contextualized} interaction.
A \emph{contextualized} interaction is \emph{explicitly purposeful} (the purpose
of the interaction, be it purely entertainment, is explicit and conscious to the
human), and takes place in an environment that fosters a stronger cognitive (and
possibly affective/social) commitment from the human in the interaction
(typically, at home). During this interaction, the human iteratively restates
and reshapes his/her behavioral and mental model of the robot (\emph{How does
the robot react to such and such situation/input?  What does the robot know
about me? About itself? About our environment? What can the robot learn?}, etc.).

@@

These cognitive levels overlap but do not exactly match the
\emph{Initialization}, \emph{Familiarization} and \emph{Stabilization} phases
introduced in our model of the dynamics of anthropomorphism. In particular,
the cognitive levels 1 and 2 are both included in the \emph{initialization} phase
of the anthropomorphism model. Sub-cognitive anthropomorphism typically
\emph{initiates} the novelty effect by rapidly engaging the human in the
interaction through an initial projected agency, whereas cognitive level 2
(projection of familiar mental models) supports the novelty effect by inducing
beliefs that the robot is set up with possibly complex cognitive abilities.

The cognitive level 3 also overlaps with the \emph{Familiarization} phase: as
the human gets used to the robot, we hypothesize one restates and adapts its
cognitive model of the robot by iteratively reshaping pre-existent, familiar
models until it provides a satisfying support to explain and justify the
observed robot behavior.

A \emph{stable level of anthropomorphism} is reached when the adaptation process
depicted in the cognitive level 3 reached a stable state, \ie the user's
experience with the robot is correctly supported by the cognitive model he/she
has built.

@@

\fixme{Discuss here the general findings of the paper: the global shape of the
model of the dynamics, ...}

@@

@@

One related question we plan to elaborate on is the effects of \emph{mutual
explicitation of the mental model of the other agent}: a human and a robot
interact on a given task, and after a while, we interrupt the task and ask each
of the agents to explicit the mental model it has built of its partner
(emotional state, beliefs, intentions, etc.). Depending on the cognitive levels
the two agents are in, we may expect varying accuracy levels in the produced
mental models.

@@

\subsection{Sustaining long-term interaction}

\paragraph{Impact of the behavioral complexity of robots}

To remain tractable, most experiments, and even more when they investigate
long-term interactions, are carried with robots exhibiting relatively simple
behaviors. In our model, the maximum level of anthropomorphism \AntMax is
reached at the end of the novelty effect, and then decrease. A natural question
is whether we can design robots such as \AntMax remains high, possibly higher
than $\ant[t_{novelty~peak}]$ (note that to know if sustaining strong
anthropomorphic interaction is desirable is entirely another question).

Intuitively, the behavioral complexity of the robot, and in particular, the ability
for the robot to evolve over time and adapt its behavior to the user as the user
familiarizes itself with the robot (\ie, \emph{co-familiarize}), seems to play a
key role.

...

@@

The article also discusses the cognitive correlates of anthropomorphism. We
proposes to identify three cognitive levels corresponding to successive
refinements of the mental models of the robot that the user builds during the
interaction. We show how these levels relate to observable anthropomorphic
effects, and how they evolve over time.
